---
title: "R Practice: Tidy Data and Joins"
format: html
---
Libraries needed!
```{r}
library(readr)
library(dplyr)
library(lubridate) # for bonus question
# I am using online links so not needed but just for reference: library(here)
```
##Pulling in the data using repo urls
```{r}
#message: false
bird_observations<-read_csv(("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=53edaa7a0e083013d9bf20322db1780e"))
bird_sites<-read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=81bf72420e69077097fb0790dcdc63a6")
bird_surveys<-read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=b2466fa5cb5ed7ee1ea91398fc291c59")
bird_taxalist<-read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=58f863b7e3066e68536a9cacdc7bd58e")
```

##Looking at Bird Observation data- 8.1 and 8.2
```{r}
#prints column names
colnames(bird_observations)
#unique values in each column
unique(bird_observations$species_id)
head(bird_observations)

```
Foreign Keys= Survey ID, Site ID, Species ID
No primary keys
##Looking at Bird Sites data
```{r}
#prints column names
colnames(bird_sites)

head(bird_sites)
```
Foreign Key= Site ID

##Looking at Bird Survey data
```{r}
#prints column names
colnames(bird_surveys)

head(bird_surveys)
```
Primary Key= Survey ID
Foriegn Key= Site ID

##Looking at Bird Taxalist data
```{r}
#prints column names
colnames(bird_taxalist)

head(bird_taxalist)
```
Primary Key= Species ID

###8.3
Write code to create a subset of bird_observations called birds_subset that only contains observations for birds with species id BHCO and RWBL, and from sites with site ID LI-W and NU-C.

```{r}
birds_subset<- bird_observations %>% filter(species_id=="BHCO"|species_id=="RWBL") %>% filter(site_id=="LI-W"|site_id=="NU-C")
head(birds_subset)

```

##8.4
#First, answer: what do you expect the outcome data frame when doing left_join() between birds_subset and sites to look like? What observations do you expect in the outcome data frame.

I expect the outcome data frame to have a lot of NAs in park district, point code, point location, and park acreage because these areas are pretty bare in bird_site dataset.There will be less species because we are only tracking two now and the site information will also be limited to two sites.

#Use a left join to update birds_subset so that it also includes sites information. For each join, include an explicit argument saying which key you are joining by (even if it will just assume the correct one for you). Store the updated data frame as birds_left. Make sure to look at the output - is what it contains consistent with what you expected it to contain?

```{r}
birds_left<- left_join(birds_subset,bird_sites, by="site_id")
head(birds_left)
```

##8.5
#what do you expect a full_join() between birds_subset and sites to contain? 
ok this one will contain all the NAs I was talking about in previous answer. 

```{r}
birds_full<-full_join(birds_subset,bird_sites,by="site_id")
head(birds_full)
```

##8.6
#what do you expect an inner_join() between birds_subset and taxalist to contain?
No NAs here I am pretty sure. This one will have the names of the birds yay! what are BHCO and RWBL red winged black bird and brown headed cowbird.

```{r}
birds_innercircle<-inner_join(birds_subset,bird_taxalist,by="species_id")
head(birds_innercircle)
```

```{r}
birds_leftcircle<-left_join(birds_subset,bird_taxalist,by="species_id")
head(birds_leftcircle)
```
idk these look the same to me sorry

##8.7 IT's Wrangling Time Sister

```{r}
bird_observations<-bird_observations %>% rename(bird_obs_notes=notes)

head(bird_observations)
```
#Combine time
```{r}
combining2<- left_join(bird_observations,bird_sites,by="site_id")
combining3<-left_join(combining2,bird_surveys,by="survey_id")
combiningitALL<-left_join(combining3,bird_taxalist,by="species_id")
head(combiningitALL)
```
#Cutting it downnn
```{r}
Cuttingitdown<-combiningitALL %>% select(survey_date,common_name,park_name,bird_count,observer)
head(Cuttingitdown)
```

##8.8

```{r}
unique(Cuttingitdown$observer)
```
6

```{r}
Cuttingitdown<- Cuttingitdown %>% mutate(observer=if_else(condition= observer == "Josh Burns", true= "J. Burns", false= observer))

unique(Cuttingitdown$observer)
```


##Cleaning Bird Observation data
```{r}

```

##Cleaning Bird Sites data
```{r}

```

##Cleaning Bird Survey data
```{r}

```

##Cleaning Bird Taxalist data
```{r}

```